---
title: "README"
author: "Alex Pinto"
date: January 12, 2015
output:
  md_document:
    variant: markdown_github
---

tiq is Functions for the Threat Intelligence IQ Tests. These should be used for the different evaluations to be run on the TI feeds.

The following functions are implemented:

- `agingTest` - Calculates the number of times an indicator is repeated on a feed throughout
- `extractPopulationFromTI` - Returns multiple population data.tables calculated using the sources on the "enriched" TI dataset on 'date'.
- `getAvailableDates` - Queries the data repository for the specific 'category' and 'group' of TI or population information and return the dates we have available for that group.
- `getDirPath` -  Builds the path for the intel analysts sources and returns it.
- `loadData` -  Lower level function helper for loadTI and loadPopulation. Should not be used directly
- `loadPopulation` -  Fetches the data contained in the Population dataset contained in 'pop.group', with the 'pop.id' as the key on the 'date' specified.
- `loadTI` -  Fetches the data contained in the Threat Intelligence dataset contained in 'category' and 'group' on the 'date' specified.
- `noveltyTest` - The novelty test is about how many indicators are being added and removed on each passing day on a specific TI feed.
- `overlapTest` - The overlap test is about how many indicators are present and repeted on multiple TI feeds.
- `plotAgingTest` - Plots an Aging Test histogram and density plot
- `plotNoveltyTest` - Plots the results of the novelty test in a (mostly) clear graph for comparisons
- `plotOverlapTest` - Plots the results of the overlap test in a (mostly) clear heatmap for comparisons
- `plotPopulationBars` -  Plots a population bar chart for simple comparisons
- `populationInference` - Runs an inference-based comparison between the reference population ('ref.pop') and the test population ('test.pop'), based on the 'pop.id' key.
- `setRootPath` - Set data path


### News

- Version 1.0 released

### Installation

```{r eval=FALSE}
devtools::install_github("mlsecproject/tiq")
```

```{r echo=FALSE, message=FALSE, warning=FALSE, error=FALSE}
options(width=120)
```

### Usage

```{r eval=FALSE}
library(tiq)

# current verison
packageVersion("tiq")

tiqtest.dir = "/SOME/DATA/DIR/tiq-test-Summer2014"
current.dir = setwd(tiqtest.dir)

tiq.data.setRootPath(file.path(current.dir, "data"))

tiq.data.getAvailableDates("raw", "public_outbound")

outbound.ti = tiq.data.loadTI("raw", "public_outbound", "20140701")

outbound.ti[, list(entity, type, direction, source, date)]
inbound.novelty = tiq.test.noveltyTest("public_inbound", "20140615", "20140630",
                                       select.sources=c("alienvault", "blocklistde",
                                                        "dshield", "charleshaley"))

tiq.test.plotNoveltyTest(inbound.novelty)

overlap = tiq.test.overlapTest("public_inbound", "20140715", "enriched",
                               select.sources=NULL)

overlap.plot = tiq.test.plotOverlapTest(overlap, title="Overlap Test - Inbound Data - 20140715")
print(overlap.plot)


outbound.pop = tiq.test.extractPopulationFromTI("public_outbound", "country",
                                                date = "20140711",
                                                select.sources=NULL, split.ti=F)
inbound.pop = tiq.test.extractPopulationFromTI("public_inbound", "country",
                                               date = "20140711",
                                               select.sources=NULL, split.ti=F)

complete.pop = tiq.data.loadPopulation("mmgeo", "country")
tiq.test.plotPopulationBars(c(inbound.pop, outbound.pop, complete.pop), "country")

```

### Test Results

```{r}
library(tiq)
library(testthat)

date()

test_dir("tests/")
```

