---
title: "Measuring the IQ of your Threat Intelligence - Summer 2014"
author: "Alex Pinto & Kyle Maxwell"
date: "August 5, 2014"
output: html_document
---

This is the companion R Markdown document to the following presentations that
were delivered in Summer 2014:

* BSides LV 2014 - "Measuring the IQ of your threat intelligence feeds"
* DEF CON 22 - "Measuring the IQ of your threat intelligence feeds"

The slides deck for these talks is located [here](http://bitly.com/tiqtest)

It should provide enough examples for usage of the tools implemented at TIQ-test.
Please review our [github repository page](https://github.com/mlsecproject/tiq-test),
report bugs and suggest features!

Adding the TIQ-TEST functions
```{r, message=FALSE}
source("tiq-test.R")
```

## Acessing the data using TIQ-TEST

We can use the `tiq.data` functions to load the Threat Intelligence datasets from
the database for exploration using R. We have defaulted to the use of `data.table`
objects for this because they are faster and you can write tighter code around it
(sorry Hadleyverse fans).

We have roughly a month of data available on this public dataset:
```{r, message=FALSE}
.tiq.data.getAvailableDates("raw", "public_outbound")
```

This is an example of "RAW" (not enriched) outbound data imported from combine output
```{r, message=FALSE}
outbound.ti = tiq.data.loadTI("raw", "public_outbound", "20140701")
outbound.ti[, list(entity, type, direction, source, date)]
```

This specific outbound dataset has the following sources included:

```{r, message=FALSE}
outbound.ti = tiq.data.loadTI("raw", "public_outbound", "20140701")
unique(outbound.ti$source)
```

We can do the same for the inbound data we have to see the sources we have available:
```{r, message=FALSE}
inbound.ti = tiq.data.loadTI("raw", "public_inbound", "20140701")
unique(inbound.ti$source)
```

SIDE NOTE: please don't add non-malicious domains to malware domain lists, ok? :)

```{r, message=FALSE}
outbound.ti = tiq.data.loadTI("raw", "public_outbound", "20140701")
outbound.ti[entity %like% "google.com", list(entity, type, direction, source, date)] 
```

We can use the same `loadTI` function to also gather the enriched datasets:
```{r, message=FALSE}
enrich.ti = tiq.data.loadTI("enriched", "public_outbound", "20140710")
enrich.ti = enrich.ti[, notes := NULL]
enrich.ti[c(2,22264, 22266)]
```

## Novelty Test

The novelty test should be used to try get a sense of the ratio of new indicators
and the retiring of old ones as the data feeds progresses day-by-day.

There is no intrinsic right or wrong, but less frequent updates usually means 
that they are carefully curated (or abandoned :) ). Curated is great, abandoned
is very bad.

Here are some results of running the Novelty test on the inbound data:

```{r, fig.height=10, fig.width=10, fig.align='center'}
inbound.novelty = tiq.test.noveltyTest("public_inbound", "20140615", "20140715", 
                                select.sources=c("alienvault", "blocklistde", 
                                                 "dshield", "charleshaley"))
tiq.test.plotNoveltyTest(inbound.novelty)
```

And results running on the outbound date:

```{r, fig.height=10, fig.width=10, fig.align='center'}
outbound.novelty = tiq.test.noveltyTest("public_outbound", "20140615", "20140715", 
                                select.sources=c("alienvault", "malwaregroup", 
                                                 "malwaredomainlist", "malwaredomains"))
tiq.test.plotNoveltyTest(outbound.novelty)
```

## Overlap Test

This is an example of applying the Overlap Test to our inbound dataset
```{r, fig.height=10, fig.width=10, fig.align='center'}
  overlap = tiq.test.overlapTest("public_inbound", "20140715", "enriched", 
                                 select.sources=NULL)
  overlap.plot = tiq.test.plotOverlapTest(overlap, title="Overlap Test - Inbound Data - 20140715")
  print(overlap.plot)
```

Similarly, an example applying the Overlap Test to the outbound dataset
```{r, fig.height=10, fig.width=10, fig.align='center'}
  overlap = tiq.test.overlapTest("public_outbound", "20140715", "enriched", 
                                 select.sources=NULL)
  overlap.plot = tiq.test.plotOverlapTest(overlap, title="Overlap Test - Outbound Data - 20140715")
  print(overlap.plot)
```

What about that day when `malwaredomainlist` and `malwaredomains` moved together on the novelty test?
```{r, fig.height=10, fig.width=10, fig.align='center'}
  overlap = tiq.test.overlapTest("public_outbound", "20140629", "enriched", 
                                 select.sources=c("alienvault", "malwaredomainlist",
                                                  "malwaredomains", "zeus"))
  overlap.plot = tiq.test.plotOverlapTest(overlap, title="Overlap Test - Outbound Data Sources - 20140629")
  print(overlap.plot)
```

## Population Test Plots

With the population data we can generate some plot to compare the top quantities
of reported IP addresses on a specific date by Country

```{r, fig.height=10, fig.width=10, fig.align='center'}
  outbound.pop = tiq.test.extractPopulationFromTI("public_outbound", "country", 
                                                  date = "20140711",
                                                  select.sources=NULL, split.ti=F)
  inbound.pop = tiq.test.extractPopulationFromTI("public_inbound", "country", 
                                                 date = "20140711",
                                                 select.sources=NULL, split.ti=F)

  complete.pop = tiq.data.loadPopulation("mmgeo", "country")
  tiq.test.plotPopulationBars(c(inbound.pop, outbound.pop, complete.pop), "country")
```

Or we can compare them by the AS that those IP addresses are a part of. Of course,
there is an infinite number of more AS's then Countries, so the distribution is much more
granular.

```{r, fig.height=10, fig.width=10, fig.align='center'}

  outbound.pop = tiq.test.extractPopulationFromTI("public_outbound", 
                                                  c("asnumber", "asname"), 
                                                  date = "20140711",
                                                  select.sources=NULL, split.ti=F)
  inbound.pop = tiq.test.extractPopulationFromTI("public_inbound", 
                                                 c("asnumber", "asname"), 
                                                 date = "20140711",
                                                 select.sources=NULL, split.ti=F)

  complete.pop = tiq.data.loadPopulation("mmasn", c("asnumber", "asname"))
  tiq.test.plotPopulationBars(c(inbound.pop, outbound.pop, complete.pop), "asname")
```

## Population Test Inference - Country data

We can use some inference tools to get a better understanding if the volume of
maliciousness we are seeing makes sense in relation to the population we consider
to be our reference population.

```{r}
outbound.pop = tiq.test.extractPopulationFromTI("public_outbound", "country", 
                                                date = "20140711",
                                                select.sources=NULL,
                                                split.ti=FALSE)
complete.pop = tiq.data.loadPopulation("mmgeo", "country")
tests = tiq.test.populationInference(complete.pop$mmgeo, 
                                     outbound.pop$public_outbound, "country",
                                     exact = TRUE, top=10)

# Whose proportion is bigger than it should be?
tests[p.value < 0.05/10 & conf.int.end > 0][order(conf.int.end, decreasing=T)]

# Whose is smaller?
tests[p.value < 0.05/10 & conf.int.start < 0][order(conf.int.start, decreasing=F)]

# And whose is the same? ¯\_(ツ)_/¯
tests[p.value > 0.05/10]
```

This tool also enables us to do trend comparison between the same TI groupings 
from different days or between different groupings. A suggested usage is comparing
the threat intelligence feeds you have against the population of confirmed attacks
or firewall blocks you have in your environment.

```{r}
outbound.pop2 = tiq.test.extractPopulationFromTI("public_outbound", "country", 
                                                 date = "20140712",
                                                 select.sources=NULL,
                                                 split.ti=FALSE)
tests = tiq.test.populationInference(outbound.pop$public_outbound, 
                                     outbound.pop2$public_outbound, "country",
                                     exact = F, top=10)

# Whose proportion is bigger than it should be?
tests[p.value < 0.05/10 & conf.int.end > 0][order(conf.int.end, decreasing=T)]

# Whose is smaller?
tests[p.value < 0.05/10 & conf.int.start < 0][order(conf.int.start, decreasing=F)]

# And whose is the same? ¯\_(ツ)_/¯
tests[p.value > 0.05/10]
```

## Population Test Inference - ASN data

We can do the same population-like tests for ASN data. Let's investigate the
prevalence of Google IP addresses on 

```{r}
outbound.pop = tiq.test.extractPopulationFromTI("public_outbound", 
                                                c("asnumber", "asname"), 
                                                date="20140711",
                                                select.sources=NULL,
                                                split.ti=FALSE)
complete.pop = tiq.data.loadPopulation("mmasn", c("asnumber", "asname"))
tests = tiq.test.populationInference(complete.pop$mmasn,
                                     outbound.pop$public_outbound, 
                                     c("asname", "asnumber"),
                                     exact = TRUE, top=10)

# Whose proportion is bigger than it should be?
tests[p.value < 0.05/10 & conf.int.end > 0][order(conf.int.end, decreasing=T)]

# Whose is smaller?
tests[p.value < 0.05/10 & conf.int.start < 0][order(conf.int.start, decreasing=F)]

# And whose is the same? ¯\_(ツ)_/¯
tests[p.value > 0.05/10]
```

This huge prevalence of Google AS IPs should be investigated further. Some of it
could be from parking at `8.8.8.8` and `1.1.1.1` but it seems to be too much

```{r, message=FALSE}
outbound.ti = tiq.data.loadTI("enriched", "public_outbound", "20140711")
outbound.ti[asname %like% "Google",  list(entity, type, source, asname, host)]
outbound.ti[asname %like% "Google" & entity != "8.8.8.8" & entity != "1.1.1.1",
            list(entity, type, source, asname, host)]
```

I guess it is fair to say that it would be a good idea to cleanup these feeds. :)

That's all for now, folks! Feel free to suggest new tests and experiments!
